---
title: "HW 3"
output:
  word_document: default
  html_document: default
---

## R Markdown

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
load("C:/Users/manya/Downloads/usgdp.rda")
library(TSA)
library(fpp)
library(tseries)
library(ggplot2)
library(forecast)
library(MLmetrics)
```


##1. Split the data into train and tests sets
```{r split, echo =TRUE}
traind<- window(usgdp, 1947, c(2005,1)) 
testd<- window(usgdp, c(2005,2), c(2006,1)) 
print(traind)

```

##2. Plot the training dataset.
```{r plot train, echo =TRUE}
autoplot(traind)
```

There doesn't seem to be a very clear seasonal pattern in the data so no transformation is needed to remove the variance's dependence on mean. However, overall there does seem to be a slight exponential trend in the data, so a transformation can be used to make the trend linear. 

##3. Plot 1st and 2nd order differences
```{r difference orders, echo =TRUE}
fstorddiff<-diff(traind)
sndorddiff<-diff(traind, differences = 2)

plot(diff(fstorddiff))
plot(diff(sndorddiff))
```

##Apply KPSS Test for Stationarity
```{r KPSS test, echo =TRUE}
kpss.test(fstorddiff)
kpss.test(sndorddiff)
```
The KPSS test for the 1st order difference results in a p-value small enough to reject the null. Because the null hypothesis for the KPSS test is stationarity, the 1st difference order results in a non-stationary dataset.The KPSS test for the 2nd order difference results in a p-value too large to reject the null. Thus the 2nd difference order results in a stationary dataset.

##4. Fit ARIMA model
```{r arima, echo =TRUE}
#Find best Lambda value and transform the data
bestLambda <- BoxCox.lambda(traind)
train_l<-BoxCox(traind, bestLambda)

#Fit ARIMA
fit <- auto.arima(train_l,seasonal=FALSE)
fit
```
This appears to be an AR Model will coefficients of 0.2829 and 0.1163, a b("drift") coefficient of 0.1851.In this model, p=2, d=1, and q=0.

##5. Compute EACF
```{r eacf, echo =TRUE}
eacf(train_l, ar.max = 2, ma.max = 2)

#Try other ARIMA models
arima_00 <-Arima(train_l, order=c(0,1,0))
arima_01 <-Arima(train_l, order=c(0,1,1))
arima_02 <-Arima(train_l, order=c(0,1,2))
arima_10 <-Arima(train_l, order=c(1,1,0))
arima_11 <-Arima(train_l, order=c(1,1,1))

#Compare AIC using summary()
summary(fit)
summary(arima_00)
summary(arima_01)
summary(arima_02)
summary(arima_10)
summary(arima_11)
```
Comparing the AIC values of the different ARIMA models, we find that auto.Arima model produces the smalles AIC value, -101.19.

##6. Forecast and plot GDP for test period
```{r forecast, echo =TRUE}
plot(forecast(fit,h=10),include=100)

```

##7. Compare forecasted values to actual test values
```{r compare_forecast, echo =TRUE}
forcast<-forecast(fit,h=4)
forc_mean<-forcast$mean
inv_forc<-InvBoxCox(forc_mean, bestLambda, biasadj = FALSE, fvar = NULL)
print("Forecast Values")
print(inv_forc)

print("Test Values")
testd 

#error
error = inv_forc - testd
print("error")
error
```

##8. Calculate MSE
```{r sse, echo =TRUE}
MSE(y_pred = inv_forc, y_true = testd)
```